---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# whatarelief

The goal of whatarelief is to build a set of raster tiles from a topography raster based on polygons. 

BEWARE: this is purely a proof of concept, not functional yet. These functions will blithely download data and use your disk space in messy ways ...

## Installation

You can install the development version of whatarelief from [Github](https://github.com/mdsumner/whatarelief) with:

``` r
devtools::install_github("mdsumner/whatarelief")
```

## Example

This is a basic example which shows you how to solve a common problem:

https://ropensci.org/blog/2018/11/13/antarctic/


```{r example}
library(blueant)
src <-   sources("ETOPO2 bathymetry")

library(whatarelief)
result <- get_relief(src)

topo <- raster::raster(find_relief())

library(ozmaps)
polys <- ozmaps::ozmap_data("states")

library(sf)
x <- relief_patcher(as(polys, "Spatial"), topo, pname = "first_successful_run")

library(raster)
purrr::walk(x, plot)

```


  Please note that the 'whatarelief' project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.
  
  
