---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![R-CMD-check](https://github.com/mdsumner/whatarelief/workflows/R-CMD-check/badge.svg)](https://github.com/mdsumner/whatarelief/actions)
<!-- badges: end -->

# whatarelief

The goal of whatarelief is to obtain elevation data. 

## Installation

You can install the development version of whatarelief from [Github](https://github.com/mdsumner/whatarelief) with:

``` r
remotes::install_github("mdsumner/whatarelief")
```

## Get elevation data

Whole planet by default, specify and extent to hone this, use 'dimension' and 'projection' for more custom options. 

```{r elevation}
library(whatarelief)
image(elevation())
image(x0 <- elevation(extent = c(120, 160, 30, 50)), col = hcl.colors(24))
contour(x0, add = TRUE, levels = 10)


image(elevation(extent = c(120, 160, 30, 50), dimension = dev.size("px")))

image(elevation(extent = c(120, 160, -50, -20), dimension = c(60, 85), resample = "near"))


image(elevation(extent = c(-1, 1, -1, 1) * 5e6, projection = "+proj=lcc +lon_0=-85 +lat_0=-42 +lat_1=0 +lat_2=-30"))
```

Can use a raster object. 

```{r rast}
elevation(terra::rast())

template <- terra::rast(terra::ext(c(-1, 1, -1, 1) * 5e6), crs = "+proj=lcc +lon_0=-85 +lat_0=-42 +lat_1=0 +lat_2=-30", ncols = 1024, nrows = 1024)
r <- elevation(template)
r

terra::plot(r)
```


Works down to quite high resolution. 

```{r resolution}
pt <- c(151.2093, -33.8688)
ex <- c(-1, 1, -1, 1) * 0.01 + rep(pt, each = 2L)
elev <- elevation(extent = ex)
image(elev, zlim = c(0, max(elev)))

```

Similar example, but a bit more context so we can easily see that it's "correct", by respatializing! 

```{r correct}
ex <- c(-1, 1, -1, 1) * 0.1 + rep(pt, each = 2L)
elev <- elevation(extent = ex)
image(elev, zlim = c(0, max(elev)))
contour(elev, levels = 5, add = TRUE)
```

We can easily "re-spatialize", but in this package we simply keep things as simple as possible.  

```{r spatial}
library(terra)
template <- terra::rast(terra::ext(ex), ncols = dim(elev)[1L], nrows = dim(elev)[2L], crs = "OGC:CRS84")
plot(setValues(template, elev[, ncol(elev):1]))
plot(sf::st_cast(ozmaps::abs_ced, "MULTILINESTRING"), add = TRUE, col = "black")
```
  
Note that all the information we need is present, we provide the extent, and the result has a dimension. (Longitude/latitude is assumed if sensible, else there's a warning - but the code will run what you ask of it). Provide all the information extent, dimension, projection to get a controlled result, or use a raster or terra object to store these. 



## Custom sources

We can provide our own sources of elevation. 

Use the 'source' argument to `elevation()`, you can input multiple sources so that a higher resolution one has
a fallback to a lower resolution one, note that by default we have "GEBCO 2019", then "NASADEM" in the sources. It doesn't matter what projection or extent these have, but usually a higher resolution one should be listed after lower. 

For example, this string provides the Amazon compute S3 elevation tiles, an 'XYZ' image server that has geotiff elements but is not a georeferenced online service (hence this small bit of XML to wrap it up). 

```{r aws-custom}
aws <- "<GDAL_WMS><Service name=\"TMS\"><ServerUrl>https://s3.amazonaws.com/elevation-tiles-prod/geotiff/${z}/${x}/${y}.tif</ServerUrl></Service><DataWindow><UpperLeftX>-20037508.34</UpperLeftX><UpperLeftY>20037508.34</UpperLeftY><LowerRightX>20037508.34</LowerRightX><LowerRightY>-20037508.34</LowerRightY><TileLevel>14</TileLevel><TileCountX>1</TileCountX><TileCountY>1</TileCountY><YOrigin>top</YOrigin></DataWindow><Projection>EPSG:3857</Projection><BlockSizeX>512</BlockSizeX><BlockSizeY>512</BlockSizeY><BandsCount>1</BandsCount><DataType>Int16</DataType><ZeroBlockHttpCodes>403,404</ZeroBlockHttpCodes><DataValues><NoData>-32768</NoData></DataValues><Cache/></GDAL_WMS>"

```

Using this, we can obtain a global summary of the available data. Sadly, these data are not quite global. 

```{r aws-global}
m <- elevation(source = aws)
image(seq(-180, 180, length.out = nrow(m)), seq(-90, 90, length.out = ncol(m)), m, asp = 1)
maps::map(add = TRUE)
abline(h = c(-90, 90), lwd = 2)

image(elevation(extent = c(-1, 1, -1, 1) * 1e7, projection = "+proj=laea +lat_0=-90", source = aws))
```

But, we can fall back to our GEBCO 2019 source to fill the gap even with a custom source (note we can provide any number of sources, in any mix of extents and projections and resolutions). 


```{r fallback}
gebco <- "/vsicurl/https://public.services.aad.gov.au/datasets/science/GEBCO_2019_GEOTIFF/GEBCO_2019.tif"
m <- elevation(source = c(gebco, aws))
image(seq(-180, 180, length.out = nrow(m)), seq(-90, 90, length.out = ncol(m)), m, asp = 1)
maps::map(add = TRUE)
abline(h = c(-90, 90), lwd = 2)

image(elevation(extent = c(-1, 1, -1, 1) * 1e7, projection = "+proj=laea +lat_0=-90", source = c(gebco, aws)))

```


## Properties of the sources in use

See vignette [elevation-sources](https://mdsumner.github.io/whatarelief/articles/elevation-sources.html). 

Note that, we could use any raster data of any kind here as custom 'source's ... (we're figuring out how to frame this package in general terms, that aren't too "spatial"). 









## Experimental features

We are working out some ways of accessing data used by the raadtools project. 

Please don't rely on them. 

```{r nsidc}
library(whatarelief)
files <- raad_source("nsidc_25km_seaice")
names(files)
idx <- which.max(files$date)

mat <- elevation(source = unlist(files[idx, c("north_vrt_dsn", "south_vrt_dsn")]))
brks <- quantile(mat[mat <= 250 & mat > 0], seq(0, 1, length.out = 16))
image(mat, col = grey.colors(length(brks) - 1), breaks = brks, main = files$date[idx])

```








## Code of Conduct
  
Please note that the whatarelief project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

